---
title: "Project 4"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
source_code: embed
---


```{r global, include=FALSE}
library(flexdashboard)
library(wesanderson)
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(here)
library(lubridate)
library(dplyr)

if (!file.exists("data")){
  dir.create(here("data"))
}

if(!file.exists(here("data","tuesdata_chocolate.rds"))){
  tuesdata <- tidytuesdayR::tt_load('2022-01-18')
  chocolate <- tuesdata$chocolate
  
  # save the files to rds objects
  saveRDS(tuesdata$chocolate, file= here("data","tuesdata_chocolate.rds"))
} else{
  chocolate <- readRDS(here("data","tuesdata_chocolate.rds"))
}

chocolate <- chocolate %>% 
  filter(!is.na(ingredients))
chocolate$cocoa_percent <- map(chocolate$cocoa_percent, function(x) str_sub(x,1,nchar(x)-1))
chocolate$cocoa_percent <- map_int(chocolate$cocoa_percent, function(x) as.integer(x))
chocolate <- chocolate %>%
  mutate(country_of_bean_origin = str_to_upper(country_of_bean_origin))
mean_ratings <- chocolate %>% 
  group_by(review_date) %>%
  summarize(mean_rating = mean(rating))
total_ratings <- chocolate %>% 
  group_by(review_date) %>%
  summarize(total_rating = n())

```


About
===

This tidytuesday dataset is about chocolate bar ratings. I want to analyze the distribution of the ratings over time and factors that affect ratings. The data and dictionary are available here: <https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md>

If you cannot view the embedded video, it is also accessible here: https://youtu.be/zX1fTT3yM8w

```{r}

library(vembedr)
embed_url("https://youtu.be/zX1fTT3yM8w") %>%
  div(class = "vembedr") %>%
  div(align = "center")

```

The Data
===

```{r, echo=FALSE}
# Create the DT table first
DT_table <- chocolate %>%
  dplyr::select(company_manufacturer,
                company_location,
                review_date,
                country_of_bean_origin,
                cocoa_percent,
                most_memorable_characteristics,
                rating)
# Instead of depending on the st_jitter algorithm to generate random placement, a custom function placing the points side by side at a set distance could be used to make points occuring at the same location appear neatly apart.
```

```{r, echo=FALSE}
DT::renderDataTable({
  DT::datatable(DT_table,
                caption = htmltools::tags$caption(
                  style = 'caption-side: top; text-align: Left;',
                  htmltools::withTags(
                    div(HTML('<a href="https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md">Click here to be redirected to a page where this data can be downloaded.</a>')))),
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '450px'))
})
```


Static: Maps
===

Column {.tabset}
-------------------------------------
   
### Bean Origin

```{r}

#adapted from https://statsandr.com/blog/world-map-of-visited-countries-in-r/

library(highcharter)
library(dplyr)
library(maps)

dat <- iso3166
dat <- rename(dat, "iso-a3" = a3)

bean_origin_countries <- unique(tolower(chocolate$country_of_bean_origin))
#change some to mapname
bean_origin_countries[bean_origin_countries == "u.s.a."] <- "usa" 
bean_origin_countries[bean_origin_countries == "sao tome"] <- "sao tome and principe"
bean_origin_countries[bean_origin_countries == "congo"] <- "republic of congo"
bean_origin_countries[bean_origin_countries == "st. lucia"] <- "saint lucia"
bean_origin_countries[bean_origin_countries == "st. vincent-grenadines"] <- "saint vincent"
bean_origin_countries[bean_origin_countries == "sao tome & principe"] <- "sao tome and principe"
bean_origin_countries[bean_origin_countries == "dr congo"] <- "democratic republic of the congo"

#match on lower case 
dat$mapname <- tolower(dat$mapname)
dat$visited <- ifelse(dat$`mapname` %in% bean_origin_countries, 1, 0)

hcmap(
  map = "custom/world-highres3", # high resolution world map
  data = dat, # name of dataset
  joinBy = "iso-a3",
  value = "visited",
  showInLegend = FALSE, # hide legend
  nullColor = "#DADADA",
  download_map_data = TRUE
) %>%
  hc_mapNavigation(enabled = FALSE) %>%
  hc_legend("none") %>%
  hc_title(text = "Countries of Bean Origin") # title

dat <- subset(dat, dat$visited == 1)
sort(dat$ISOname) # sort is to have the visited countries in alphabetical order



```   
 
### Company Location

```{r}

#adapted from https://statsandr.com/blog/world-map-of-visited-countries-in-r/

library(highcharter)
library(dplyr)
library(maps)

dat <- iso3166
dat <- rename(dat, "iso-a3" = a3)

company_countries <- unique(tolower(chocolate$company_location))
#change some to mapname
company_countries[company_countries == "u.s.a."] <- "usa" 
company_countries[company_countries == "u.k."] <- "uk(?!r)" 
company_countries[company_countries == "sao tome"] <- "sao tome and principe"
company_countries[company_countries == "st. lucia"] <- "saint lucia"
company_countries[company_countries == "st. vincent-grenadines"] <- "saint vincent"
company_countries[company_countries == "sao tome & principe"] <- "sao tome and principe"
company_countries[company_countries == "u.a.e."] <- "united arab emirates"

#match on lower case 
dat$mapname <- tolower(dat$mapname)
dat$visited <- ifelse(dat$`mapname` %in% company_countries, 1, 0)

hcmap(
  map = "custom/world-highres3", # high resolution world map
  data = dat, # name of dataset
  joinBy = "iso-a3",
  value = "visited",
  showInLegend = FALSE, # hide legend
  nullColor = "#DADADA",
  download_map_data = TRUE
) %>%
  hc_mapNavigation(enabled = FALSE) %>%
  hc_legend("none") %>%
  hc_title(text = "Countries of Companies") # title

dat <- subset(dat, dat$visited == 1)
sort(dat$ISOname) # sort is to have the visited countries in alphabetical order

```   

Static: Cocoa Percent
===

```{r}

mean_cocoa <- chocolate %>% 
  group_by(review_date) %>%
  summarize(mean_rating = mean(cocoa_percent))


ggplot(data = mean_cocoa, aes(x=review_date, y=mean_rating)) +
  geom_point() +
  ylim(c(65,75)) +
  labs(title = "Mean Cocoa Percent over Time", subtitle = "cocoa content stays relatively constant around 71%") +
  xlab("Year") + ylab("Mean Cocoa Percent") +
  theme_bw()

```

Interactive: Cocoa Percent
===

Column {.sidebar}
-----------------------------------------------------------------------

Distribution of Cocoa Percent for this dataset

```{r}
selectInput("n_breaks", label = "Number of bins:",
            choices = c(4, 6, 8, 10), selected = 5)

```

Column
-----------------------------------------------------------------------

### Distribution of Cocoa Percent

```{r}
renderPlot({
  hist(chocolate$cocoa_percent, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Cocoa Percent", main = "Distribution of Cocoa Percent")

})
```

Interactive: Boxplots
===

Column {.sidebar}
-----------------------------------------------------------------------

To see how rating varies by different factors

```{r}
selectInput("variable", "Variable:",
                  c("Year of Review" = "review_date",
                    "Cocoa Percent" = "cocoa_percent"))
checkboxInput("outliers", "Show outliers", TRUE)
```

Column
-----------------------------------------------------------------------


```{r}
formulaText <- reactive({
    paste("rating ~", input$variable)
  })

renderPlot({
    boxplot(as.formula(formulaText()),
            data = chocolate,
            outline = input$outliers,
            col = "#75AADB", pch = 19, main=formulaText())
  })
```


Analysis
===

Column
-----------------------------------------------------------------------


```{r, eval=TRUE, message=FALSE}
  
ggplot(data = mean_ratings, aes(x=review_date, y=mean_rating)) +
  geom_point() +
  ylim(c(0,5)) +
  labs(title = "Mean Rating over Time", subtitle = "ratings stay relatively constant around 3-3.5", caption = "Figure 1") +
  xlab("Year") + ylab("Mean Rating") +
  theme_bw()
```

Column
-----------------------------------------------------------------------


```{r, eval=TRUE, message=FALSE}
  
pal <- wes_palette("Royal1")
ggplot(data = total_ratings, aes(x=review_date, y=total_rating)) +
  geom_point() +
  geom_line(color=rep(pal,4)) +
  ylim(c(0,300)) +
  labs(title = "Total Ratings per Year over Time", subtitle = "ratings fluctuate over time, peaking in 2015", caption = "Figure 2") +
  xlab("Year") + ylab("Number of Ratings") +
  theme_bw()
```

Column
-----------------------------------------------------------------------


```{r, eval=TRUE, message=FALSE}
pal <- wes_palette("Darjeeling1")
ggplot(chocolate, aes(x = rating, fill = review_date)) + 
  geom_density() + 
  facet_wrap(~ review_date) + 
  theme_bw() +
  labs(title = "Rating Density over Time", subtitle = "this plot shows the variation in reviews across years", caption = "Figure 3") +
  xlab("Rating") + ylab("Density") +
  theme(legend.position="none",
      strip.background=element_rect(color=pal[5],
                                    fill=pal[3]))
```



